---
title: "Data Visualization Exercises"
output: html_notebook
---

Using `mpg` dataset
```{r}
library(ggplot2)
library(ggthemes)
head(mpg)
colnames(mpg)

```


## Histogram of hwy mpg values:
```{r}
mpgHist <- ggplot(mpg, aes(x=hwy)) + geom_histogram(binwidth=1, aes(fill=model))
mpgHist
```

 
## Barplot of car counts per manufacturer with color fill defined by cyl count
```{r}
carBar <- ggplot(mpg, aes(x=manufacturer)) + geom_bar(aes(fill=factor(cyl))) 
carBar
```


## Switch now to use the txhousing dataset that comes with ggplot2
```{r}
head(txhousing)
colnames(txhousing)
```


## Create a scatterplot of volume versus sales. Afterwards play around with alpha and color arguments to clarify information.
```{r}
#carPlot <- carPlot + geom_point(aes(shape=factor(cyl),color=factor(cyl)))

# change the color scale
#carPlot <- carPlot + geom_point(aes(color=hp)) + scale_color_gradient(low='blue',high='red')

volSales <- ggplot(txhousing, aes(x=volume,y=sales)) + geom_point(aes(color=factor(month)), size=2, alpha=0.4) 
 
volSales

```

## Add a smooth fit line to the scatterplot from above. Hint: You may need to look up geom_smooth()
```{r}
smoothVolSales <- volSales + geom_smooth(color='red')
smoothVolSales

```

