---
title: "K Means Clustering Project"
output: html_notebook
---
 
Usually when dealing with an unsupervised learning problem, its difficult to get a good measure of how well the model performed. For this project, we will use data from the UCI archive based off of red and white wines (this is a very commonly used data set in ML).

We will then add a label to the a combined data set, we'll bring this label back later to see how well we can cluster the wine into groups.



Use read.csv to open both data sets and set them as df1 and df2. Pay attention to what the separator (sep) is. 
 
 
# imports and get data
```{r}
library(readr)
library(ggplot2)
library(cluster)
library(dplyr)

df1 <- read.csv('Training Exercises/Machine Learning Projects/CSV files for ML Projects/winequality-red.csv', sep = ";")

df2 <- read.csv('Training Exercises/Machine Learning Projects/CSV files for ML Projects/winequality-white.csv', sep = ";")




```

## Now add a label column to both df1 and df2 indicating a label 'red' or 'white'.
```{r}
df1$Label <- "Red"
df2$Label <- "White"
```


## inspect the data
```{r}
head(df1)
```
 

```{r}
head(df2)
```

## Combine df1 and df2 into a single data frame called wine.
```{r}
wine <- df1 %>% full_join(df2)
head(wine)
tail(wine)
```

```{r}
str(wine)
```

```{r}
summary(wine)
```

 
# EDA

## Create a Histogram of residual sugar from the wine data. Color by red and white wines.
```{r}
ggplot(wine, aes(x=residual.sugar)) + geom_histogram(aes(fill=Label), col='black', bins = 50) + scale_fill_manual(values = c('#8B0000','#F6FEB1'))
```

## Create a Histogram of citric.acid from the wine data. Color by red and white wines.
```{r}
ggplot(wine, aes(x=citric.acid)) + geom_histogram(aes(fill=Label), col='black', bins = 50) + scale_fill_manual(values = c('#8B0000','#F6FEB1'))
```

## Create a Histogram of alcohol from the wine data. Color by red and white wines.
```{r}
ggplot(wine, aes(x=alcohol)) + geom_histogram(aes(fill=Label), col='black', bins = 50) + scale_fill_manual(values = c('#8B0000','#F6FEB1'))
```

## Create a scatterplot of residual.sugar versus citric.acid, color by red and white wine.
```{r}
ggplot(wine, aes(x=residual.sugar, y=citric.acid)) + geom_point(aes(col=Label, shape=Label), alpha=0.4) + scale_color_manual(values = c('#8B0000','#F6FEB1')) + theme_dark()

```

## Create a scatterplot of volatile.acidity versus residual.sugar, color by red and white wine.
```{r}
ggplot(wine, aes(x=residual.sugar, y=volatile.acidity)) + geom_point(aes(col=Label, shape=Label), alpha=0.4) + scale_color_manual(values = c('#8B0000','#F6FEB1')) + theme_dark()
```

# Clustering

## Grab the wine data without the label and call it clus.data
```{r}
# stop at the 12th column because we want to get rid of Label.
clus.data <- wine[,1:12]
head(clus.data)
```


## calculate the best K number using elbow method
This isn't necessary (we only have two clusters), but I just found out how to do it so I'll leave the code in the project.
```{r}
library(factoextra)
fviz_nbclust(clus.data, kmeans, method= "wss") 
# + geom_vline(xintercept = 3, linetype = 2)
```



## Call the kmeans function on clus.data and assign the results to wine.cluster.
```{r}
set.seed(101)

# arguments for kmeans: dataset, K, number of iterations
wine.cluster <- kmeans(clus.data, 2, nstart = 20)

wine.cluster
```

## Print out the wine.cluster Cluster Means and explore the information.
```{r}
wine.cluster$centers
```

# Evaluating the Clusters
You usually won't have the luxury of labeled data with KMeans, but let's go ahead and see how we did!

## Use the table() function to compare your cluster results to the real results. Which is easier to correctly group, red or white wines?
```{r}
table(wine$Label, wine.cluster$cluster)

```
Red is easier to cluster. 



