---
title: "K-Nearest Neighbour Practice"
output: html_notebook
---

In this notebook I practise the K-nearest neighbour method, using the `Default` dataset from the ISLR library. 

The question I want to answer is:

Can we identify whether a given person will default?

## 1. Get the data
```{r}
library(ISLR)
head(Credit)
```

### 2. Clean the data
The `student` variable is non-numeric, which could be a problem. I don't think converting it to 0/1 would work (would it? let's try)

```{r}

  
```




### 3. Standardise the data
Use scale() to standardize the feature columns of the dataset. Set this standardized version of the data as a new variable. _Note that because the variables in this dataset are of a fairly similar order of magnitude, there's no need to standardise - I'm doing it for practice.

Check that the scaling worked by checking the variance of one of the new columns.


### 4. Create training and test data

Use the caTools library to split your standardized data into train and test sets. Use a 70/30 split.





### 5. Build the `knn` model




### 6. Identify the best k number

