---
title: "Linear Progression Project"
output: html_notebook
---


 
## import libraries and csv
```{r}
library(readr)
library(ggplot2)
library(plotly)
library(ggthemes)
library(dplyr)

bike <- read.csv('Training Exercises/Machine Learning Projects/CSV files for ML Projects/bikeshare.csv')
head(bike)
tail(bike)
summary(bike)

numBike <- sapply(bike,is.numeric)
cor(bike[numBike])

```

## Can you figure out what is the target we are trying to predict? 
The target would be the count variable. That is, the number of bicycles used. This would be useful for the people who run a bike rental scheme - they would want to know what the demand is for their product at different times, in different places, so they can purchase the right number of bicycles.

 
## Exploratory Data Analysis
Create a scatter plot of count vs temp. Set a good alpha value.
```{r}
ggplot(bike, aes(x=temp, y=count, color=temp)) + geom_point(alpha=0.4)
```


## Plot count versus datetime as a scatterplot with a color gradient based on temperature. You'll need to convert the datetime column into POSIXct before plotting.
```{r}
bike <- bike %>% mutate(datetime=as.POSIXct(datetime))
#head(bike)
ggplot(bike, aes(x=datetime, y=count, color=temp)) + geom_point(alpha=0.4)
```

What is the correlation between temp and count?
```{r}
cor(bike$temp, bike$count)
```

Let's explore the season data. Create a boxplot, with the y axis indicating count and the x axis begin a box for each season.
```{r}
ggplot(bike, aes(x=factor(season), y=count)) + geom_boxplot(aes(fill=factor(season)),alpha=0.5) + coord_flip() + theme_classic()



# use coord_flip to flip the axes
# carBox <- ggplot(carData,aes(x=factor(cyl),y=mpg)) + geom_boxplot(aes(fill=factor(cyl))) + coord_flip() + theme_linedraw()
# carBox

```



## Feature Engineering
A lot of times you'll need to use domain knowledge and experience to engineer and create new features. Let's go ahead and engineer some new features from the datetime column.

Create an "hour" column that takes the hour from the datetime column. You'll probably need to apply some function to the entire datetime column and reassign it. Hint:

time.stamp <- bike$datetime[4]
format(time.stamp, "%H")

```{r}
bike <- bike %>% mutate(hour=format(datetime, "%H"))
head(bike)
```


## Now create a scatterplot of count versus hour, with color scale based on temp. Only use bike data where workingday==1.
```{r}
bike2 <- bike %>% filter(workingday==1)

myColors <- c('red', 'green', 'yellow', 'pink', 'blue')

ggplot(bike2, aes(x=hour, y=count,color=temp)) + geom_point(alpha=0.4, position = position_jitter(w=1,h=0)) + scale_color_gradientn(colors=myColors)




```

Optional Additions:

Use the additional layer: scale_color_gradientn(colors=c('color1',color2,etc..)) where the colors argument is a vector gradient of colors you choose, not just high and low.
Use position=position_jitter(w=1, h=0) inside of geom_point() and check out what it does.

Now create the same plot for non working days:
```{r}
bike3 <- bike %>% filter(workingday==0)
ggplot(bike3, aes(x=hour, y=count,color=temp)) + geom_point(alpha=0.4, position=position_jitter(w=1,h=0)) + scale_color_gradientn(colors=myColors)
```


Building the Model
Use lm() to build a model that predicts count based solely on the temp feature, name it temp.model
```{r}
tempModel <- lm(count ~ temp, data=bike)

```


Get the summary of the temp.model
```{r}
summary(tempModel)
```

## How many bike rentals would we predict if the temperature was 25 degrees Celsius? Calculate this two ways:

Using the values we just got above
Using the predict() function
```{r}

myWay <- 6.0462+(25 * 9.1705)
myWay

temp.test <- data.frame(temp=c(25))
 
highWay <- predict(tempModel, temp.test)
highWay

```

## Use sapply() and as.numeric to change the hour column to a column of numeric values.
```{r}

bike$hour <- sapply(bike$hour,as.numeric)
head(bike)
```


## Finally build a model that attempts to predict count based off of the following features. Figure out if theres a way to not have to pass/write all these variables into the lm() function. Hint: StackOverflow or Google may be quicker than the documentation.

season
holiday
workingday
weather
temp
humidity
windspeed
hour (factor)

```{r}
model <- lm(count ~ . -casual - registered -datetime -atemp,bike )
```

Get the summary of the model
```{r}
summary(model)
```







