---
title: "Machine Learning: Linear Regression"
output: html_notebook
---
Section 20 of the course


# 1. Import and explore the data


## import and inspect data
```{r}
library(readr)

df <- read.csv('Machine Learning with R/student-mat.csv', sep = ';')
head(df)
summary(df)
```

## check if there are any NA values
```{r}
any(is.na(df))
```

## get correlations of the numeric variables
```{r}
library(ggplot2)
library(ggthemes)
library(dplyr)

numCols <- sapply(df,is.numeric)

corData <- cor(df[numCols])
corData
```
## import some packages for plotting correlations
```{r}
library(corrgram)
library(corrplot)

# corrplot requires numeric data
corrplot(corData,method='color')

```

```{r}
# corrgram is similar to corrplot, but you can use the original data without filtering for numeric values only.
corrgram(df,order=T,lower.panel = panel.pie, upper.panel = panel.shade, text.panel = panel.txt)
```

```{r}
# looking at a histogram 
ggplot(df,aes(x=G3)) + geom_histogram(bins=20,alpha=0.5,fill='blue')
```

# 2. Split the data into training and testing data

## install caTools package
```{r}
library(caTools)

# set a seeds file of random data
set.seed(101)

# state the way you are splitting the data for the sample. arguments are a dataframe column and splitRatio (the proportion of the data you want to use for the sample)
sample <- sample.split(df$G3,SplitRatio = 0.7)

# create the training subset (i.e. here it's 0.7 of the data)
train <- subset(df, sample == TRUE)

# which means the test subset should be 0.3
test <- subset(df,sample==FALSE)
```

## create the training model
```{r}

# for a linear regression model
# y is the thing you want to learn/predict, x1 x2 of data are the things that will help the model predict
#model <- lm(y ~ x1 + x2,data)

# quicker way for a linear regression model
#model <- lm(y ~ . , data)

model <- lm(G3 ~ ., data = train)
```

## run the model
```{r}
summary(model)
```
R-squared - the higher the score, the better the fit to the model (1 is maximum)


## plot the residuals
The residuals show the difference between reality and your model (so, the smaller the residual values, the more accurate your model)
Ideally, we want to see a normal distribution. Something unusual (e.g. binomial distribution) might indicate that a linear regression model is a bad strategy.

```{r}
res <- residuals(model)
class(res)
res <- as.data.frame(res)

ggplot(res,aes(res)) + geom_histogram(fill='blue', alpha=0.5)
```
Note that there are negative `res` values. This doesn't make sense, as the lowest result possible is 0 (it's a test). However, there is a way to get round this. 

```{r}
plot(model)
```

# 3. Test the predictions

```{r}
G3Predictions <- predict(model, test)
results <- cbind(G3Predictions, test$G3)
colnames(results) <- c('pred','real')
results <- as.data.frame(results)
head(results)
min(results)
```

## deal with those strange negative predictions
since the minimum score on a test is zero, transform any negative values into zero.
```{r}
toZero <- function(x) {
  if (x <0) {return(0)} else {return(x)}
}

results$pred <- sapply(results$pred,toZero)
min(results) # we can see that the minimum is now zero, so the function toZero() has worked
```

## evaluate the prediction values
```{r}
# 1. use Mean Standard Error
mse <- mean((results$real-results$pred)^2)
mse
```

```{r}
# 2. Root Mean Squared Error
mse^0.5
```

```{r}
# 3. R-squared value
SSE <- sum((results$pred-results$real)^2)
SST <- sum((mean(df$G3)-results$real)^2)
R2 <- 1- SSE/SST
R2
```
This score suggests that the model is quite a good fit (score is almost 0.8, whilst perfection would be 1)
