---
title: "Natural Language Processing 2"
output: html_notebook
---
 
I have followed a tutorial from [towards data science](https://towardsdatascience.com/create-a-word-cloud-with-r-bde3e7422e8a)

I also tried an alternative method, but didn't have success. I may have made an error, so I am going to try again.
[sthda](http://www.sthda.com/english/wiki/word-cloud-generator-in-r-one-killer-function-to-do-everything-you-need)

## import libraries
* tm for text mining
* wordcloud and wordcloud 2for generating word cloud images
* RColorBrewer for color palettes
* dplyr for manipulating and cleaning data
* I used readr to read the source text, though this library may not be necessary.
```{r}

library(tm)
library(wordcloud)
library(wordcloud2)
library(RColorBrewer)
library(readr)
library(dplyr)

```


## get data
```{r}

recommendations <- read_tsv('FATFrecommendations.txt')

```



## turn the data into a corpus
```{r}

recommendationsCorpus <- Corpus(VectorSource(recommendations))
 

```

## clean the data
```{r}

recommendationsCorpus <- recommendationsCorpus %>%
  tm_map(removeNumbers) %>%
  tm_map(removePunctuation) %>%
  tm_map(stripWhitespace)
recommendationsCorpus <- tm_map(recommendationsCorpus, content_transformer(tolower))
recommendationsCorpus <- tm_map(recommendationsCorpus, removeWords, stopwords("english"))


```

## create a document-term-matrix
```{r}

dtm <- TermDocumentMatrix(recommendationsCorpus)
matrix <- as.matrix(dtm)
words <- sort(rowSums(matrix), decreasing = T)
df <- data.frame(word = names(words), freq=words)

```

## make the word cloud
```{r}

set.seed(1234) # for reproducibility 
wordcloud(words = df$word, freq = df$freq, min.freq = 100, max.words=200, scale=c(2.7,0.23), random.order=FALSE, rot.per=0.35, colors=brewer.pal(8, "Dark2"))

```

## make a word cloud using the `wordcloud2` library, which has additional features.

```{r}

wordcloud2(data=df, size=1.6, color='random-dark')

```

## change the shape of the cloud
```{r}

cloud2 <- wordcloud2(data=df, size = 0.7, shape = 'star')
cloud2
```


## save a word cloud to .png (could also do pdf)
Note that it would default to html, so we need `webshot` and `htmlwidgets`.

This process is very slow: it takes several minutes.
```{r}

library(webshot)
webshot::install_phantomjs(force = T)
library(htmlwidgets)

saveWidget(cloud2, "cloud2.html", selfcontained = F)

webshot("cloud2.html", "cloud2.png", delay = 5)



# webshot("cloud2.html","cloud2.png", delay =5, vwidth = 480, vheight=480)


```


