---
title: "Section 15: Data Visualization"
output: html_notebook
---

The `ggplot2` library uses the concept of __layers__. The layers are:
* Data
* Aesthetics
* Geometries
* Facets - means we can put multiple plots on a single canvas
* Statistics - e.g. you can add a line of best fit around a scatter plot
* Coordinates - you can limit x and y axes
* theme - e.g. font size, colours

Data, aesthetics and geometries are the core layers.
 
## import the library 
```{r}
library(ggplot2)
```


## histograms
```{r}
library('ggplot2movies')

# create data layer and aesthetics
movieGram <- ggplot(movies, aes(x=rating))

# add geometries layer
# alpha affects the transparency of the plot
#movieGram <- movieGram + geom_histogram(binwidth=0.1, color='red', fill='pink', alpha=0.4)

movieGram <- movieGram + geom_histogram(binwidth=0.1, aes(fill=..count..))

# add labels
movieGram <- movieGram + xlab('Movie Rating') + ylab('Count')

# add title
movieGram <- movieGram + ggtitle('My Movie Ratings')

print(movieGram)

```

##scatterplots
```{r}
df <- mtcars

# data and aesthetics layer
carPlot <- ggplot(df, aes(x=wt,y=mpg))

# geometry
# base the size on hp by putting it into aesthetics
#carPlot <- carPlot + geom_point(aes(size=hp))

# use factor if you want to have a different shape and colour for each type of categorical variable
#carPlot <- carPlot + geom_point(aes(shape=factor(cyl),color=factor(cyl)))

# change the color scale
carPlot <- carPlot + geom_point(aes(color=hp)) + scale_color_gradient(low='blue',high='red')

carPlot
```

## bar charts
```{r}
# import the mpg dataset that comes with ggplot2
df <- mpg

# create the data and aesthetics
mpgBar <- ggplot(df,aes(x=class))

#add geometry

#basic
#mpgBar <- mpgBar + geom_bar()


# split the bars by another variable (drv)
#mpgBar <- mpgBar + geom_bar(aes(fill=drv))

# subdivide the bars by another variable (uses position = 'dodge' to create two separate bars per class of vehicle)
#mpgBar <- mpgBar + geom_bar(aes(fill=drv),position = 'dodge')

# use position = 'fill' to show how each class is composed- the percentage of each drv within a class 
mpgBar <- mpgBar + geom_bar(aes(fill=drv),position = 'fill')


mpgBar
```

## boxplots
```{r}
carData <- mtcars

# use coord_flip to flip the axes
carBox <- ggplot(carData,aes(x=factor(cyl),y=mpg)) + geom_boxplot(aes(fill=factor(cyl))) + coord_flip() + theme_linedraw()
carBox
```

## 2-variable plotting
```{r}
movieData <- movies

# moviePlot <- ggplot(movieData, aes(x=year,y=rating)) + geom_bin2d(binwidth=c(3,1)) + scale_fill_gradient(high = 'red',low = 'blue')

# moviePlot <- ggplot(movieData, aes(x=year,y=rating)) + geom_hex() + scale_fill_gradient(high = 'red',low = 'blue')

moviePlot <- ggplot(movieData, aes(x=year,y=rating)) + geom_density2d() 



moviePlot
```

## co-ordinates 
```{r}

plotto <- ggplot(mpg, aes(x=displ,y=hwy)) + geom_point()


plotto2 <- plotto + coord_cartesian(xlim=c(1,4),ylim=c(15,30))

# set aspect ratio
plotto3 <- plotto + coord_fixed(ratio = 1/3)

plotto
plotto2
plotto3

```


## faceting
```{r}

# the drv is what you want to facet by on x axis. the cyl is what you want to facet by on the y axis. Instead of a column, you can use a full stop.
#facetPlot <- plotto + facet_grid(drv ~ cyl)
#facetPlot <- plotto + facet_grid(. ~ cyl)
facetPlot <- plotto + facet_grid(drv ~ .)
facetPlot
```

## themes
```{r}

# you can set the theme for all subsequent themes
#theme_set(theme_minimal())

 
plotty <- ggplot(mtcars,aes(x=wt,y=mpg)) + geom_point()

# you can add a theme to a specific plot
#plotty <- plotty + theme_dark()

# using themes from ggthemes package
plotty <- plotty + theme_pander()
plotty

```


