---
title: "Section 8: Matrices"
output: html_notebook
---

 
## create a vector using a range
```{r}
deci <- 1:10
deci
```


## turn `deci` into a matrix
```{r}
matrix(deci,nrow = 5)
```

## populate a matrix by row
```{r}
rower <- matrix(1:12, byrow = FALSE, nrow = 4)
rower
```

## Create a matrix from two vectors, and give column and row names
```{r}
lena <- c(9, 6, 7, 8, 6)
cama <- c(8, 7, 5, 9, 6)

scores <- c(lena, cama)
scores.matrix <- matrix(scores, nrow = 2)


years <- c("2015", "2016", "2017", "2018", "2019")
names <- c("Lena", "Cama")
colnames(scores.matrix) <- years
rownames(scores.matrix) <- names
scores.matrix
```

## Arithmetic with matrices
```{r}
# imagine that the system for allocating scores has changed. The new system has double the points. So, we will multiply all the scores in scores.matrix by 2.
scores.matrix * 2
```

```{r}
# sum the columns and add a new row
annualTotal <- colSums(scores.matrix)

#scoresPlusSums <- scores.matrix + sums
#scoresPlusSums <- cbind(scores.matrix, row_sums)
scoresPlusSums <- rbind(scores.matrix, annualTotal)


lifeTimeTotal <- rowSums(scoresPlusSums)
scoresPlusSums <- cbind(scoresPlusSums, lifeTimeTotal)
scoresPlusSums
```

```{r}

pentathlon <- 1:10
pentamatrix <- matrix(pentathlon, nrow = 5)
pentamatrix
#pentamatrix[1,]
print('---')
pentamatrix[1:3,2]


```

```{r}
# Imagine that one of our players has cheated. They add a point to all their scores.
scoresPlusSums['Lena',] <- scoresPlusSums['Lena',] + 1
scoresPlusSums
```

```{r}
# The judges have noticed a problem with the scores. They're not sure what's going on, but they decide to give Cama extra points in 2019 so she can win. 
scoresPlusSums['Cama', '2019'] <- scoresPlusSums['Cama', '2019'] + 5
scoresPlusSums
# Interesting... The lifeTimeTotal column for Cama did not increase. 
```

```{r}
animal <- c('d', 'c', 'd', 'c', 'c')
id <- c(1,2,3,4,5)
factor(animal)
temps <- c('cold', 'med', 'hot', 'hot', 'hot', 'cold', 'med')
fact.temp <- factor(temps, ordered = T,levels = c('cold', 'med', 'hot'))
fact.temp
summary(fact.temp)

```

Ex 1: Create 2 vectors A and B, where A is (1,2,3) and B is (4,5,6). With these vectors, use the cbind() or rbind() function to create a 2 by 3 matrix from the vectors. You'll need to figure out which of these binding functions is the correct choice.
```{r}
A <- c(1,2,3)
B <- c(4,5,6)
#vec1 <- matrix(A, nrow = 1)
#vec1 <- rbind(vec1, B)
  

vec <- rbind(A,B)
vec

```


Ex 2: Create a 3 by 3 matrix consisting of the numbers 1-9. Create this matrix using the shortcut 1:9 and by specifying the nrow argument in the matrix() function call. Assign this matrix to the variable mat
```{r}
mat <- matrix(1:9, nrow = 3)
mat
```



Ex 3: Confirm that mat is a matrix using is.matrix()
```{r}
is.matrix(mat)
```



Ex 4: Create a 5 by 5 matrix consisting of the numbers 1-25 and assign it to the variable mat2. The top row should be the numbers 1-5.
```{r}
mat2 <- matrix(1:25, byrow = T, nrow = 5)
mat2
```



Ex 5: Using indexing notation, grab a sub-section of mat2 from the previous exercise that looks like this:
[7,8]
[12,13]
```{r}
mat2[2:3,2:3]
```



Ex 6: Using indexing notation, grab a sub-section of mat2 from the previous exercise that looks like this:
[19,20]
[24,25]
```{r}
mat2[4:5,4:5]
```




Ex 7: What is the sum of all the elements in mat2?
```{r}
sum(mat2)
```



Ex 8: Ok time for our last exercise! Find out how to use runif() to create a 4 by 5 matrix consisting of 20 random numbers (4*5=20).
```{r}
#help("runif")
mat3 <- matrix(runif(20, 1, 1000), nrow = 4 )
mat3
```

