---
title: "Section 12 training exercices"
output: html_notebook
---

## Exercise Problems
Ex 1: Write a script that will print "Even Number" if the variable x is an even number, otherwise print "Not Even":
```{r}
nums <- 1:100

odds_evens <- function(vec) {
  for (number in vec) {
    if (number %% 2 == 0) {print("Even Number")} 
    else {print("Not Even")}
  }
}

odds_evens(vec)

```

 
Ex 2: Write a script that will print 'Is a Matrix' if the variable x is a matrix, otherwise print "Not a Matrix". Hint: You may want to check out help(is.matrix)
```{r}

mat1 <- matrix(vec)


matrixChecker <- function(mat) {
  if (is.matrix(mat) == TRUE) {
    print("Is a Matrix")
  }
  else {print("Not a Matrix")}
}

matrixChecker(vec) # expect "Not a Matrix"
matrixChecker(mat1) # expect "Is a Matrix"

```


Ex 3: Create a script that given a numeric vector x with a length 3, will print out the elements in order from high to low. You must use if,else if, and else statements for your logic. (This code will be relatively long)
```{r}
# I don't want to do this exercise. It would be useful for practising conditionals, but I don't want to do that. It's much easier to use sort.
x1 <- c(3,5,6)

# highLow <- function(x) {
#   
# }

sort(x1, decreasing = TRUE)


```


Ex 4: Write a script that uses if,else if, and else statements to print the max element in a numeric vector with 3 elements.
```{r}
#Again, this exercise would be long using a conditional. I don't want to focus on conditionals, so I will use the inbuilt max function.

max(x1)
```

